---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Market basket analysis

Ref: https://edu.kpfu.ru/pluginfile.php/278552/mod_resource/content/1/MachineLearningR__Brett_Lantz.pdf

- Using simple performance measures to find associations in large databases
- Understanding the peculiarity of transactional data
- Knowing how to identify the useful and actionabe patterns

The building block of a market analysis are the items that may appear in any given transaction.

## Association rules

### Principle

Association rules are always composed by subsets of itemsets and are denoted by relating one of the itemset on the left-hand sile (LHS) of the rule to another itemset on the right-hand side (RHS). The LHS is the condition that needs to be met in order to trigger the rule, and the RHS is the expected result of meeting that condition.

{peanut butter, jelly} --> bread

This association rule states that if peanut butter and jelly are purchased together, then bread is also likely to be purchased. In other words, "peanut butter and jelly imply bread".

In this context, association rules are not used for prediction, but rather for unsupervised knowledge discovery in large databases. The program is simply unleashed on a dataset in the hope that interesting associations are found? The downside is that there isn't an easy way to objectively measure the performance of a rule learner, aside from evaluating it for qualitative usefulness.

Association rules are also helpful for finding patterns in many different types of data. Other applications include:

- Searching for interesting and frequently occurring patterns of DNA and protein sequences in cancer data
- Finding patterns of purchases or medical claims that occur in combination with fraudulent credit card or insurance use
- Identify combinations of behavior that precede customers dropping their cellular phone service or upgrading their cable television package 

Association rule analysis is used to search for interesting connections among very large number of elements.

### Measures

- Support:

This says how popular an itemset is, as measured by the proportion of transactions in which an itemset appears

support = count(X)/N

where N is the number of transactions and count(X) is the number of transactions containing itemset X.

- Confidence:

It is a measurement of its predictive power or accuracy. It is defined as the support of the itemset containing both X and Y divided by the support of the itemset containing only X

confidence(X -> Y) = support(X,Y)/support(X)

## The Apriori algorithm

The most widely-used approach for efficiently searching large databases for rules is known as Apriori. The algorithm utilizes a simple prior belief about the properties of frequent itemsets: all subsets of a frequent itemset must also be frequent.

Strengths:
- Capable of working with large amount of transactional data
- Results in rules that are easy to understand
- Useful for data mining and discovering unexpected knowledge in databases

Weaknesses:
- Not very helpful for small datasets
- Takes effort to separate the true insight from the common sense
- Easy to draw spurious conclusions from random patterns


## Apply in

### Install packages 

```{r}
#install.packages("arules")
library(arules)
```

### Data preparation

#### Summary statistics

```{r}
data(Groceries)
head(itemInfo(Groceries))
```
```{r}
summary(Groceries)
```


Number of rows refer to the number of transactions and the number of columns indicate each of the different items.

Each cell in the matrix is a 1 if the item was purchased for the corresponding transaction, or 0 otherwise.

The density value refers to the proportion of non-zero matrix cells.

We can deduct then that the total number of items = (number of rows * number of columns * density) and the average transaction = ((number of rows * number of columns)/number of rows) distinct grocery items

The next block of summary() outputs lists the items that were commonly found in the transactional data. Since 2,513/9,835 (number of rows), we can determine that whole milk appeared in 25.6 % of transactions.

Finally, the other set of statistics are about the size of the transactions. A total of 2159 transactions contained only a single item, while one transaction had 32 items. The first quartile and median purchase size are two and three items, implying that 25 % of the transactions contained two or fewer items and about half contained three items or fewer.


#### Other arules package features 

To look at the contents of the sparse matrix:

```{r}
# the first five transactions
inspect(Groceries[1:5])
```


To examine a particular item (a column of data), use [row, column].  The itemFrequency() function allows to see the proportion of transactions that contain the specified item.

To view the support level for the first three items:

```{r}
itemFrequency(Groceries[,1:3])
```


#### Visualizing item support - item frequencyplots

FrequencyPlot() function creates a bart chart depicting the proportion of transactions containing specified items. 
it is possible to only display items that appear in a minimum proportion of transactions:

```{r}
#Histogram showing the 8 items with at least 10% support
itemFrequencyPlot(Groceries, support=0.1)
```


The support levels for the top 20 grocery items is shown below. The histogram is sorted by decreasing support.
```{r}
# Histogram limiting the plot to a specific number of items
itemFrequencyPlot(Groceries, topN=20)
```












